---
title: "Investigation of theoretical mean and quantile"
author: "Abubakar HARUNA"
date: "04/02/2022"
output: 
  html_document: 
    fig_caption: yes
    fig_height: 3.5
    fig_width: 4
    toc: yes
    keep_md: true
---

```{r setup, include=FALSE}
library(knitr)
library(viridisLite)
library(ismev)

library(data.table)

library(ggplot2)
knitr::opts_chunk$set(echo = F)
opts_chunk$set(warning = F)
opts_chunk$set(message = F)
```

# Part 1



```{r}
#load the data
data = read.csv2(file = "../Data/lausanne.csv", header = T, sep = ";", stringsAsFactors = F, )
summary(data)
str(data)
data$tre200nx = as.numeric(data$tre200nx)
data$time = as.Date(as.character(data$time), "%Y%m%d")

summary(data)
#plot(data$time, data$tre200nx)


```

(a) Plot the yearly maxima. What do you observe ?

```{r}
#convert to yearly data maximum

year_vec = lubridate::year(data$time)
yearly_data =aggregate(data$tre200nx, list(year_vec), max, na.rm =T)
colnames(yearly_data) = c("year", "obs")

plot(yearly_data$year, yearly_data$obs)



```



(b) Fit a constant-parameter GEV distribution to the yearly maxima and assess your model using quantile and probability plots. Comment. Discuss the estimated value of the shape parameter.

```{r}
#?ismev
mod_0=gev.fit(yearly_data$obs)
gev.diag(mod_0)
fit_gev$mle
#comment/ negative shape param


```


(c) Fit non-constant-parameter GEV distributions to the yearly maxima and discuss the significance of a yearly varying location/shape/scale parameter.

```{r}
mod_1 =gev.fit(xdat = yearly_data$obs,ydat = matrix(yearly_data$year, ncol = 1 ),  mul = 1)
mod_2 =gev.fit(xdat = yearly_data$obs,ydat = matrix(yearly_data$year, ncol = 1 ),  shl  = 1)
mod_3 =gev.fit(xdat = yearly_data$obs,ydat = matrix(yearly_data$year, ncol = 1 ),  sigl  = 1)

AIC(mod_0, mod_1, mod_2, mod_3)
```

(d) Using the selected model calculate the probability that next's year maximum will exceed all
previous levels.
(e) Estimate the 10-year return level from this model (Practical 2).